### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@dad08218838f5926c9b134a7afbd8009ad71a986 🚀

Spillover risk is not unidimensional. From the standpoint of an animal
community, \emph{i.e.} a pool of possible hosts, it is driven by a
multiplicity of factors (Plowright et al. 2017). The global richness of
hosts is one such component commonly mentioned/analysed (see \emph{e.g.}
Anthony et al. 2017 for coronaviruses), but there is an argument to be
made that species who are not competent hosts of a specific virus genus
may not factor into this (Plowright et al. 2015), or that species who
are assumed to share viruses at different rates should be weighted
accordingly (Albery et al. 2020). In mammals, key functional traits (for
which phylogeny is a reasonable proxy) are determinants of the spillover
potential (Olival et al. 2017); these include, notably, body mass and
affinity for urban environments (Albery et al. 2022). Finally,
especially when the pool of potential hosts spans the entire globe,
there may be local host pools that are highly unique; not having been
observed in other locations, these can act on the overall risk either by
providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid changes in specialism (Agosta et al. 2010). In the specific case
of generalist pathogens (as betacoronavirus clearly are), there is
conceptual and empirical support to the idea that these community- level
mechanisms are even more important in driving the overall risk (Power
and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
threaten human health. Especially concerning is the fact that bats are
reservoirs for a variety of emerging viruses (Calisher 2006), making
balancing the needs for bat conservation and disease prevention a
potentially difficult act, especially in more densely populated areas
(REF). Chiropterans emerged around 64 million years ago and are one of
the most diverse mammalian orders, with an estimated richness of more
than 12000 species, (Peixoto F et al, 2018) and 14325 known species
Simmons \& Cirranello. They exhibit a broad variety of habitat use,
behaviour, and feeding strategies, resulting in their playing an
essential role in the delivery of several ecosystem services (Kasso
2013), including economic benefits. Over two-thirds of bats are either
obligate or facultative insectivorous mammals, therefore playing an
important role in the regulation of insect pests that can affect crops
(Williams-Guillen 2011), and vectors of diseases that put a risk on
human health (Gonsalves 2013 ). Because bats are globally distributed
and have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the extant distribution of
coevolutionary processes between bats and the pathogens they carry. As a
consequence, not all areas are facing a risk of human spillover, and
those who do might not be facing risks of the same nature and magnitude.

Yet a comprehensive assessment of the risk of spillover of
betacoronaviruses from bat hosts to humans is limited by the fact that
we do not know the full diversity of viruses associated with every bat
species. Predictive models can help fill in some of these gaps, by
recommending hosts based on known host-virus associations BECKER REF. In
this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of known
and predicted hosts. We turn these associations into a spatially
explicit additive mapping of zoonotic risk components, which reveals
extreme heterogeneity of risk at the global scale; furthermore, we
identify the Amazon and South-Eastern Asia as hotspots of phylogenetic
distinctiveness of betacoronaviruses; surprisingly, current data suggest
that viral sharing between hosts in high in the Amazon and low in
South-Eastern Asia, which has the potential to result in different
evolutionary dynamics between these two regions.

TK summary of the results

>>>>

Spillover risk is not unidimensional. From the standpoint of an animal
community, \emph{i.e.} a pool of suitable hosts, it is driven by a
multiplicity of factors (Plowright et al. 2017). The global richness of
hosts is one such component commonly mentioned/analysed (see \emph{e.g.}
Anthony et al. 2017 for coronaviruses), but there is an argument to be
made that species who are not competent (or know) hosts of a specific
virus genus may not factor into this (Plowright et al. 2015), or that
species who are assumed to share viruses at different rates should be
weighted accordingly (Albery et al. 2020). In mammals, key functional
traits (for which phylogeny is a reasonable proxy) are determinants of
the spillover potential (Olival et al. 2017); these include, notably,
body mass and affinity for urban environments (Albery et al. 2022).
Finally, especially when the pool of potential hosts spans the entire
globe, there may be local host pools that are highly unique; not having
been observed in other locations, these can act on the overall risk
either by providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid evolutionary changes in specialism of their pathogens (Agosta et
al. 2010). In the specific case of generalist pathogens, there is
conceptual and empirical support to the idea that these community- level
mechanisms are even more important in driving the overall risk (Power
and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
threaten human health. Chiropterans emerged around 64 million years ago
and are one of the most diverse mammalian orders, with an estimated
richness of more than 12000 species, (Peixoto F et al, 2018) and 14325
known species (Simmons \& Cirranello). They exhibit a broad variety of
habitat use, behaviour, and feeding strategies, resulting in their
playing an essential role in the delivery of several ecosystem services
tied to important ecosystem-derived benefits (Kasso 2013). For example,
over two-thirds of bats are know to be either obligate or facultative
insectivorous mammals, therefore playing an important role in the
regulation of insect pests that can affect crops (Williams-Guillen
2011), and vectors of diseases that put a risk on human health
(Gonsalves 2013). Because bats are globally distributed and have a long
evolutionary history, phylogeographic and biogeographic approaches are
required to shed light on the extant distribution of coevolutionary
processes between bats and the pathogens they carry. Not all areas in
which bats, viruses, and human are co-occuring are facing a risk of
spillover towards human populations, and the areas in which this risk
exist may not be facing risks of the same nature and magnitude.

In this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of known
and recently discovered hosts. This work is important both as a
description of the bats-betacoronavirus complex, but aslo because more
broadly, bats are known reservoirs for a variety of emerging viruses
(Calisher 2006), making balancing the needs for bat conservation and
disease prevention a potentially difficult act and a source of
human-wildlife conflicst, especially in more densely populated areas
(REF). By drawing on concepts from the Geographic Mosaic Theory of
Coevolution (REF), we turn these associations into a spatially explicit
additive mapping of zoonotic risk components, which reveals extreme
heterogeneity of risk at the global scale; furthermore, we identify the
Amazon and South-Eastern Asia as hotspots of phylogenetic
distinctiveness of betacoronaviruses; surprisingly, current data suggest
that viral sharing between hosts in high in the Amazon and low in
South-Eastern Asia, which has the potential to result in different
evolutionary dynamics between these two regions.
